---
title: "GPT helps ..."
author: "Sebastián Tyrrell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Certainly! Below is an R Markdown document that incorporates the explanation and R script designed to identify the best predictor feature for classifying between "virginica" and "versicolor" in the `iris` dataset.

```markdown
---
title: "Identifying the Best Predictor in Iris Dataset"
output: html_document
---

```{r setup, include=FALSE}
# Load necessary libraries
library(caret)
library(tidyverse)
```

## Introduction
In this analysis, we aim to determine which feature in the `iris` dataset is the most effective predictor for distinguishing between "virginica" and "versicolor". By evaluating different cutoff values for each feature, we can identify the threshold that maximizes classification accuracy.

## Data Preparation
We begin by loading and preparing the dataset. We'll exclude the "setosa" species to focus on the two other species.

```{r data-preparation}
# Load and filter dataset
data(iris)
iris <- iris[-which(iris$Species == 'setosa'),]

# Partition data into training and test sets
set.seed(76)
y <- iris$Species
test_index <- createDataPartition(y, times = 1, p = 0.5, list = FALSE)
train <- iris[-test_index,]

# Identify feature columns (numeric features only)
features <- colnames(iris)
features <- features[-length(features)]  # Remove 'Species'
```

## Analysis Process
For each feature, we'll compute the classification accuracy over a range of cutoff values and identify the cutoff that yields the highest accuracy.

```{r analysis-process}
# Initialize a tibble to store results
results <- tibble(Feature = character(),
                  MaxAccuracy = numeric(),
                  OptimalCutoff = numeric())

# Iterate over each feature
for (i in features) {
  # Generate cutoff sequence
  cutoff <- seq(min(train[[i]], na.rm = TRUE), max(train[[i]], na.rm = TRUE), 0.1)
  
  # Calculate accuracy for each cutoff
  accuracy <- map_dbl(cutoff, function(x) {
    temp_y_hat <- ifelse(train[[i]] > x, "virginica", "versicolor") |> 
      factor(levels = levels(train$Species))
    mean(temp_y_hat == train$Species)
  })
  
  # Find maximum accuracy and corresponding cutoff
  max_accuracy <- max(accuracy)
  optimal_cutoff <- cutoff[which.max(accuracy)]
  
  # Store results
  results <- results %>% add_row(Feature = i, MaxAccuracy = max_accuracy, OptimalCutoff = optimal_cutoff)
}
```

## Results
We'll inspect the results and identify the feature that maximizes accuracy.

```{r results}
# Print results
print(results)

# Determine the best feature
best_feature <- results %>% filter(MaxAccuracy == max(MaxAccuracy))
print(best_feature)
```

## Conclusion
The analysis identifies which feature serves as the best predictor based on classification accuracy. Understanding the optimal cutoff for this feature can help in making more informed decisions in classification tasks involving similar datasets.

```

### How to Use
1. **Open** your R environment, such as RStudio.
2. **Create a new R Markdown file**: Copy and paste the content above into the new file.
3. **Knit** the document to an HTML file or PDF to execute the code and generate the report.

This R Markdown document will help you dynamically generate and display the results of your classification analysis in a clear and organized manner. Let me know if you need further customization!

(Sensitivity level: CU) GPT@EC AI generated content – please use with caution.